---
title: "Video's notes"
author: "Victoria"
date: "2/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ***Basic arithmetic***

### Addition
```{r}
1 + 1
```

### Subtraction
```{r}
5 - 3
```

### Division
```{r}
2/3
```

### Power
```{r}
2^3
```

### Product
```{r}
2 * 3
```

### Remainder (Modulus)
```{r}
3 %% 2
```

# ***Variables***
```{r}
variable <- 100 #to assign it
variable #to print it
bank <- 1000
bank

#This are good practices to name variables:
bank.account <-123
bankAccount <-133

#Assigning results into variables
deposit <- 100
total <- bank.account + deposit
total #printing the result

```

# ***Data types***
```{r}
1 #integer
1.5 #float or decimals
"hello" #character
TRUE #logical or boolean

class(deposit) #to check the class of a variable
```

# ***Vector***
```{r}
nvec <- c(1,2,3,4,5) #a vector is created using the "combination function" [c()], each number represents a column
nvec #we print it

cvec <- c("U","S","A")
cvec

lvec <- c(T, F)
lvec
```

### Combining data types in a vector
Rule: **never combine** different data types in a vector, or it will forcefully change them to the same data type, e.g. TRUE in a numeric vector changes to "1".
```{r}
v <- c(TRUE, 20, 40)
v
class(v)
```

### Vector names
We can assign a name to **each value in a vector**, for example, to associate days with a vector of temps values
```{r}
temps <- c(72,71,68,73,69,75,23)
temps

names(temps) <- c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")
temps #now the vector has a name for each value
```
Another way to do it, faster, is by creating the names vector independently and then use it
```{r}
days <- c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")
names(temps) <- days
temps
```

# ***Vector operations***
```{r}
v1 <- c(1,2,3)
v2 <- c(5,6,7)
v1 + v2 #It will make the operation ELEMENT BY ELEMENT
v1 * v2
v1 - v2
v1 / v2

# To sum all the elements in a numeric vector
sum(v1)
sum.of.v1 <- sum(v1)
sum.of.v1

# mean of a vector
mean(v1)

# standard deviation of a vector
sd(v1)

# maximum value
max(v1)

# minimum value
min(v1)

# product of all the elements in a vector
prod(v1)
```

# ***Comparison operators***
They will return a TRUE or FALSE according to the operation
```{r}
5 > 6
5 < 6
5 <= 6
2 == 3 #equality
2 != 4 #inequality

my.var <- 2
my.var > 5
```

When comparing vector values, it will return a **vector result** filled with TRUE and FALSE
```{r}
v <- c(1,2,3,4,5)
v < 2 # Comparison is done ELEMENT BY ELEMENT
v == 3
```

We can also compare two vectors
```{r}
v2 <- c(10, 20, 30, 40, 50)
v < v2
```

# ***Vector indexing and slicing***
NOTE: Indexing starts at number 1
```{r}
v1 <- c(100,200,300)
v2 <- c("a","b","c")
v1
v2

#Lets start indexing values from the vectors
v1[1]
v1[3]
v2[3]
```
Slicing means we're gonna take multiple items from a vector at the same time
```{r}
#Lets start slicing
v1[c(1,2)] #to grab the 1st and 2nd element
v2[c(1,3)] #to grab a and c

#to grab continuous elements
v <- c(1,2,3,4,5,6,7,8,9,10)
v
v[2:5] #To grab from the 2nd element till the 5th element of the vector
v[7:10]
```

Lets assign names to the values of our slice, and **we can actually index it as the name given instead of the number of element**
```{r}
v.slice <- v[7:10]
v.slice
v.names <- c("a","b","c","d")

names(v.slice) <- v.names #names assigned
v.slice

# We can index the values as the names given
v.slice[2]
v.slice["b"]
v.slice[c("c","a","d")]
```

To grab only values greater than 2
```{r}
v
v.labels <- c("a","b","c","d","e","f","g","h","i","j")
names(v) <- v.labels

v[v>2] # it returns only the values that are greater than 2

# Another way to do it
my.filter <- v>2
v[my.filter]
```
# ***To search for help and info about terms and funcitons in R***
```{r}
help("vector")
??vector
help.search("vector")
```

# ***....R Basics Exercises...***
What is 2 to the power of 5
```{r}
2^5
```
Create a vector called stock.prices with the following data points: 23,27,23,21,34
```{r}
stock.prices <- c(23,27,23,21,34)
```
Assign names to the price data points relating to the day of the week, starting with Mon, Tue, Wed, etc...
```{r}
stock.names <- c("Mon","Tue","Wed","Thu","Fri")
names(stock.prices) <- stock.names
stock.prices
```
What was the average (mean) stock price for the week? (You may need to reference a built-in function)
```{r}
mean(stock.prices)
```
Create a vector called over.23 consisting of logicals that correspond to the days where the stock price was more than $23
```{r}
over.23 <- stock.prices[stock.prices>23]
```

Use the over.23 vector to filter out the stock.prices vector and only return the day and prices where the price was over $23

```{r}
over.23
```

Use a built-in function to find the day the price was the highest
```{r}
max(stock.prices)
```

