---
title: "Section 8 - R Matrices"
author: "Victoria"
date: "2/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ***Creating a matrix***
To create a matrix out of a vector
```{r}
v <- 1:10 #easy way to create a vector

matrix(v) #creates a matrix with 10 rows and 1 column
```
To create a matrix from zero
```{r}
matrix(v,nrow = 2) # I want a matrix with 2 rows and whatever columns with the values of v
matrix(1:12,byrow = FALSE, nrow = 4) # it starts filling the values vertically
matrix(1:12,byrow = TRUE, nrow = 4) # it starts filling the values horizontally (row by row)
```

```{r}
goog <- c(450,451,452,445,468)
msft <- c(230,231,232,233,220)

stocks <-c(goog,msft) #to combine both vectors in one
print(stocks)

stock.matrix <- matrix(stocks, byrow = TRUE, nrow = 2)
print(stock.matrix)

#To name the rows and columns values
days <- c("Mon","Tue","Wed","Thu","Fri")
st.names <- c("GOOG","MSFT")

colnames(stock.matrix) <- days
rownames(stock.matrix) <- st.names
print(stock.matrix)
```

# ***Matrix Arithmetic***
All the operations made to the matrix are made element by element
```{r}
mat <- matrix(1:25, byrow = TRUE, nrow = 5)
mat
# Examples of operations
mat * 2
mat - 2
mat / 2
mat^2
mat > 15

#We can also filter out some values of the matrix in a form of a vector
mat.f <- mat[mat>15]
mat.f

mat * mat # element by element multiplication
mat %*% mat # real linear algebra matrix multiplication
```

# ***Matrix operations***
We start with the previous matrix
```{r}
goog <- c(450,451,452,445,468)
msft <- c(230,231,232,233,220)

stocks <-c(goog,msft) #to combine both vectors in one
print(stocks)

stock.matrix <- matrix(stocks, byrow = TRUE, nrow = 2)
print(stock.matrix)

#To name the rows and columns values
days <- c("Mon","Tue","Wed","Thu","Fri")
st.names <- c("GOOG","MSFT")

colnames(stock.matrix) <- days
rownames(stock.matrix) <- st.names
print(stock.matrix)
```
Operations on the matrix
```{r}
colSums(stock.matrix) #it adds all the values on each column
rowSums(stock.matrix) #it adds all the values on each row

rowMeans(stock.matrix) #mean value of the rows
colMeans(stock.matrix) #mean value of the columns
```
Exercise: Add a vector as a new row of your previous matrix
```{r}
Fbook <- c(111,112,113,114,115)
# Function rbind
tech.stocks <- rbind(stock.matrix,Fbook) # rbind adds the selected vector as a new row of the selected matrix
tech.stocks
# Function cbind
avg <- rowMeans(tech.stocks)
avg
tech.stocks <- cbind(tech.stocks,avg) # rbind adds the selected vector as a new column of the selected matrix
tech.stocks
```

# ***Matrix selection and Indexing***
Creating the matrix
```{r}
mat <- matrix(1:50, byrow = TRUE, nrow = 5)
mat
```
Lets index!
```{r}
mat[1,] #all the columns of the 1st row
mat[,1] #all the rows of the 1st column
mat[] #all the values of the matrix
mat[1:3,] #all the columns of the rows 1 to 3
mat[1:2,1:3]
mat[,9:10]
mat[2:3,5:6]
```

# ***Factor and categorical matrices***

**Factor function** is useful to create categorical matrices
There are 2 types of categorical variables:

- **Nominal:** There's NO order  
- **Ordinal:** There's order (like numbers, temp, letters a to z)  

Nominal example:
```{r}
animal <- c("dog","cat","dog","cat","cat")
id <- c(1,2,3,4,5)

factor(animal) # Finds the categories within the values of our vector
fact.ani <- factor(animal)

```
Ordinal example with temps:
```{r}
ord.cat <- c("cold","med","hot")
temps <- c("cold","med","hot","hot","hot","cold","med")
temps
fact.temp <- factor(temps, ordered = TRUE, levels = c("cold","med","hot")) #we can tell the factor function our levels of categorization
fact.temp
```

We can use this to find the # of times a value appears in a vector
```{r}
# Summary function
summary(fact.temp) #of the categorized vector

summary(temps) # of the uncategorized vector (not the expected results)
```

# **...R Matrix Exercises...**

Create 2 vectors A and B, where A is (1,2,3) and B is (4,5,6). With these vectors, use the cbind() or rbind() function to create a 2 by 3 matrix from the vectors. You'll need to figure out which of these binding functions is the correct choice.

```{r}
A <- 1:3
B <- 4:6
matrix.AB <- rbind(A,B)
matrix.AB
```

Create a 3 by 3 matrix consisting of the numbers 1-9. Create this matrix using the shortcut 1:9 and by specifying the nrow argument in the matrix() function call. Assign this matrix to the variable mat
```{r}
mat <- matrix(1:9, nrow = 3)
mat
```
Confirm that mat is a matrix using is.matrix()
```{r}
is.matrix(mat)
```

Create a 5 by 5 matrix consisting of the numbers 1-25 and assign it to the variable mat2. The top row should be the numbers 1-5.
```{r}
mat2 <- matrix(1:25,byrow = TRUE, nrow = 5)
mat2
```

Using indexing notation, grab a sub-section of mat2 from the previous exercise that looks like this: 
[7,8]
[12,13]
```{r}
mat2[2:3,2:3]
```
Using indexing notation, grab a sub-section of mat2 from the previous exercise that looks like this:
[19,20]
[24,25]
```{r}
mat2[4:5,4:5]
```
What is the sum of all the elements in mat2?
```{r}
sum(mat2)
```
Ok time for our last exercise! Find out how to use runif() to create a 4 by 5 matrix consisting of 20 random numbers (4*5=20).
```{r}
v <- runif(20, min = 0, max = 100) #creates a vector of random numbers from min to max
v
v.mat <- matrix(v,byrow = FALSE, nrow = 4)
v.mat
```

