---
title: "Section 13 - Advanced R Programming"
author: "Victoria"
date: "6/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**Built-in R Features and Functions**
## Seq(): Create a sequence
```{r}
seq(0,100,by=2) #creates a vector from 0 to 100, with a step-size of 2
seq(0,24,by=10)
```
## Sort(): Sort a vector
```{r}
v <- c(1,4,7,2,13,3,11)
v
sort(v) #it sorts the values in ascending order
sort(v,decreasing = TRUE) #it sorts the values in descending order
cv <- c("a","c","d")
sort(cv)
```
## rev(): Reverse elements in object
```{r}
v <- 1:10
v
rev(v)
```
## append(): Merge objects together (works on vectors and lists)
```{r}
v <- 1:10
v2 <- 35:40
append(v,v2) #it merges them together
```
## Checking the **Data type** in R
```{r}
v <- c(1,2,3)

is.vector(v) #we check if it's a vector or not
is.data.frame(mtcars)
```
## To **convert into another Data type**
```{r}
v <- 1:3

as.list(v)
as.matrix(v)
```

#**Apply**
##Sample function: to sample objects from a vector
```{r}
sample(x = 1:100,1) #we get a certain sample of random numbers from a vector x
```
##Apply function: to apply a function over all the values in a vector or list
```{r}
v <- c(1,2,3,4,5)

addrand <- function(x){
  ran <- sample(1:100,1)
  return(x+ran)
}

print(addrand(10))

result <- lapply(v,addrand) #as a list
print(result)

result <- sapply(v,addrand) #as a vector
print(result)
```
Another example
```{r}
v <- 1:5

times2 <- function(num){
  return(num*2)
}
print(v)

result <- sapply(v,times2)
print(result)
```
##Anonymous functions
```{r}
v <- 1:5

result <- sapply(v,function(num){num*2}) #we just created the anonymous function and used it
print(result)
```

```{r}
v <- 1:5

add_choice <- function(num,choice){
  return(num+choice)
}

print(sapply(v,add_choice,choice=100))
```
# **Math functions in R**
##abs():absolute value
```{r}
abs(-2)
abs(2)
v <- c(-1,-2,-3)
abs(v)
```
##sum(): the sum of a vector
```{r}
sum(v)
```
##mean(): the mean of a vector
```{r}
mean(v)
```
##round(): round values (additional arguments to nearest)
```{r}
round(2.88887,digits = 0)
```
**TIP: Google "R reference card" for the most useful cheat sheet of functions in R**

#**Regular Expressions**
##grepl: returns a logical
##grep: returns an index
```{r}
text <- "Hi there, how are you?"
text

grepl("how",text) #to check if there's that element in the variable
grepl("dog",text)
```
```{r}
v <- c("a","b","c","d")
grepl("b",v) #to find if the statement is true or false
grep("b",v) #returns the indexes in which the value appears

```

#**Date and Timestamps**
## Dates
```{r}
Sys.Date() #year-month-day

today <- Sys.Date()
class(today)
```

We normally get the input date as a character, so we need to transfrom it into a date object in the **STANDARD DATE FORMAT!**
```{r}
c <- "1990-01-01"
class(c)
c.Date <- as.Date(c)
class(c.Date)
```
sometimes we don't get the dates in the needed format
```{r}
## This will give NA(s) in some locales; setting the C locale
## as in the commented lines will overcome this on most systems.
## lct <- Sys.getlocale("LC_TIME"); Sys.setlocale("LC_TIME", "C")
x <- c("1jan1960", "2jan1960", "31mar1960", "30jul1960")
z <- as.Date(x, "%d%b%Y")
## Sys.setlocale("LC_TIME", lct)

my.date <- as.Date("Nov-03-90",format = "%b-%d-%y")
my.date

my.date1 <- as.Date("June,01,2002",format="%B,%d,%Y")
my.date1
```
**as.Date format code list:**
- %d Day (decimal)  
- %m Month (decimal)  
- %b Month (abbreviated)  
- %B Month (full name)  
- %y Year (2 digit)  
- %Y Year (4 digit)  

## **Time information**
R stores time info in POSIXct class
```{r}
#to convert strings into data type for time stamp information
as.POSIXct("11:02:03",format="%H:M:%S") #we actually don't use this, we use the following instead:

help("strptime") #here we find all the glossary of conversion from string to time info
strptime("11:02:03",format="%H:%M:%S") #now, it is a time data!!


```

