---
title: "Section 10 - R Lists"
author: "Victoria"
date: "5/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**R Lists**
Lists allows **different types of data structure** in the same variable
**NOTE:**lists is more of an organizational tool, dataframes and matrices are to work with data directly
```{r}
v <- c(1,2,3) #Vector
m <- matrix(1:10,nrow=2) #matrix
df <- mtcars #data frame

class(v)
class(m)
class(df)
```
### To create a list
```{r}
my.list <- list(v,m,df)
my.list
```
We can also assign names instead of the automatic list order
```{r}
my.named.list <- list(sample.vec = v, my.matrix = m, sample.df = df)
my.named.list

#to call certain elements of the list
my.named.list$my.matrix #to call the matrix with $
my.named.list[["my.matrix"]]
my.named.list[[1]]

class(my.named.list)

#to get the list element in its original format, use the $ method
class(my.named.list$my.matrix)
double.list <- c(my.named.list,my.named.list) #to combine 2 lists together
```
The best way to get info about the list is with the str function
```{r}
str(my.named.list)
```

# **Data Input and Output**
###**CSV Input and Output**
CSV are the most common data files to work with data. 
- The first line: indicates the columns names  
- The rest of the rows are data separated by commas
```{r}
write.csv(mtcars, file = "my_example.csv") #to create a new csv file with an existing data frame
```
To **read a csv file**
```{r}
ex <- read.csv("another_example.csv") #we create a data frame out of the csv file
ex #a data frame!
head(ex)
tail(ex)
```
To **write to a csv file**
```{r}
write.csv(ex,file = "my_new_example.csv")
read.csv("my_new_example.csv")
```

# **Excel Files with R**
```{r}
library(readxl) #to call the library
excel_sheets("Sample-Sales-Data.xlsx") #this tells us the name of the sheets that the file has
df <- read_excel("Sample-Sales-Data.xlsx", sheet = "Sheet1")
head(df) #the excel file is read as a data frame
```
Now we can make the regular **data frame operations**
```{r}
mean(df$Value)
sum(df$Value)
summary(df)
str(df)
```
##**Lets use lists to store a whole excel file with different sheets (aka data frames)**
List = whole excel workbook
Data frames = sheets
```{r}
#To read a simple xlsx file as a dataframe
library(openxlsx)

df.wb <- read.xlsx( "Sample-Sales-Data.xlsx")

#To read a complete workbook of excel as a dataframe
df.complete <- loadWorkbook("Reporte de Actividades (Ago-Dic 2020).xlsx") #to load the workbook
df.sheet1 <- read.xlsx(df.complete,sheet = 1) #to create a dataframe of each sheet
df.sheet2 <- read.xlsx(df.complete,sheet = 2)
df.sheet3 <- read.xlsx(df.complete,sheet = 3)

```
## How to write into an Excel file
```{r}

df2xl <- mtcars #we create an example data frame to export as an excel file
write.xlsx(df2xl,file = "output_example1")
```
To create a workbook of how many excel sheets we want
```{r}
## setup a workbook with 3 worksheets
wb <- createWorkbook()
addWorksheet(wb = wb, sheetName = "Sheet 1", gridLines = FALSE)
writeDataTable(wb = wb, sheet = 1, x = iris) #x is the dataframe of info to go on that sheet
addWorksheet(wb = wb, sheetName = "mtcars (Sheet 2)", gridLines = FALSE)
writeData(wb = wb, sheet = 2, x = mtcars)
addWorksheet(wb = wb, sheetName = "Sheet 3", gridLines = FALSE)
writeData(wb = wb, sheet = 3, x = Formaldehyde)
worksheetOrder(wb)
names(wb)
worksheetOrder(wb) <- c(1, 3, 2) # switch position of sheets 2 & 3
writeData(wb, 2, 'This is still the "mtcars" worksheet', startCol = 15)
worksheetOrder(wb)
names(wb) ## ordering within workbook is not changed

```

#**SQL with R**
SQL is the most used programming languaged used to interact with databases online
**There are special packages for each database** we have to search in Google: database + R
```{r}
#install.packages("RODBC")
#RODBC Example of syntax
#library(RODBC)
```
another example
```{r}
library(RPostgreSQL)
```
And finally google in R-bloggers for examples of the required library

#**Web Scrapping with R**
You need to know HTML, CSS and the %<% R operation to do manual web scrapping (obtaining csv files from websites) **OR** you can enter to [www.import.io] and paste the URL to get the csv file of data
## **Manual method**
```{r}
#demo(package = "rvest")
#demo(package = "rvest",topic = "tripadvisor") #we'll get a built in tutorial to use it
```

