---
title: "Section 15 - Data Visualization with R"
author: "Victoria"
date: "7/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**Overview of ggplot2**
##ggplot2
Here is the cheat sheet for each kind of plot: {https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf}
Is the most popular data visual package for R
It consists of 3 main layers:
- Geometries  
- Aesthetics  
- Data  
And 3 optional layers:
- Facets  
- Statistics  
- Coordinates
```{r}
library(ggplot2) #library
pl <- ggplot(data = mtcars,aes(x=mpg, y=hp)) #data and aesthetics
pl + geom_point() #geometry
# DONE!

```
Optional layers
##Facets: allows us to put multiple plots in a single canvas
```{r}
pl <- ggplot(data = mtcars,aes(x=mpg, y=hp)) + geom_point() #data, aesthetics & geometry
pl + facet_grid(cyl ~ .) #facets, each cyl value has its own plot!
```

##Statistics: allows us to have a shadow of error (stadistical value)
```{r}
pl <- ggplot(data = mtcars,aes(x=mpg, y=hp)) + geom_point() #data, aesthetics & geometry
pl + facet_grid(cyl ~ .) + stat_smooth()#facets, each cyl value has its own plot!

```
##Coordinates: allows us to set limits of the x and y axis
```{r}
pl <- ggplot(data = mtcars,aes(x=mpg, y=hp)) + geom_point() #data, aesthetics & geometry
pl2 <- pl + facet_grid(cyl ~ .) + stat_smooth()#facets, each cyl value has its own plot!
pl2 + coord_cartesian(xlim = c(15,25))
```
#Theme: to affect the details like fonts, lines, etc.
```{r}
pl <- ggplot(data = mtcars,aes(x=mpg, y=hp)) + geom_point() #data, aesthetics & geometry
pl2 <- pl + facet_grid(cyl ~ .) + stat_smooth()#facets, each cyl value has its own plot!
pl2 + coord_cartesian(xlim = c(15,25)) + theme_dark()
```

#**Histograms**
They are used to plot the **frecuency** of a value in our data frame
```{r}
library(ggplot2)
library(ggplot2movies)

#Data & Aesthetics layers
pl <- ggplot(movies,aes(x=rating))
#Geometry
pl2 <- pl + geom_histogram(binwidth = 0.1,color="red",fill="pink",alpha=0.4) #the binwidth is the "width" of the values for each block
#alpha controls the transparecy of our blocks
pl3 <- pl2 + xlab("Movie Rating") + ylab("Count")

print(pl3 + ggtitle("Quantity of movie ratings"))
```

#**Scatterplots**
To show correlations between 2 variables
```{r}
df <- mtcars
df
# Data & Aesthetics layer
pl <- ggplot(df,aes(x=wt,y=mpg))
# Geometry layer
pl.op1 <- pl + geom_point(size=2,alpha=0.5)
pl.op2<- pl + geom_point(aes(shape=factor(cyl),color=factor(cyl)),size=5) #the size of the points change according to cyl values (this dependency of format and values only works by using aes())

print(pl.op2)
```
##Color
We can use HEX color arguments too! You can google the HEX code of your favorite color and just paste it as color (#code of the color)
```{r}
df <- mtcars
# Data & Aesthetics layer
pl <- ggplot(df,aes(x=wt,y=mpg))
# Geometry layer
pl.op2<- pl + geom_point(aes(shape=factor(cyl)),size=5,color="#56ea29") #the size of the points change according to cyl values (this dependency of format and values only works by using aes())

print(pl.op2)
```

##Scale color gradient
We can set the gradient of colors we want for the scatterplot
```{r}
df <- mtcars
# Data & Aesthetics layer
pl <- ggplot(df,aes(x=wt,y=mpg))
# Geometry layer
pl1<- pl + geom_point(aes(color=hp),size=5)

pl2 <- pl1 + scale_color_gradient(low = "blue",high = "red")

print(pl2)
```

#**Barplots**
It's similar to a histogram, but we deal with **categorical data** on the x axis, not with data, the y axis is a count too.
```{r}
df <- mpg

pl <- ggplot(df,aes(x=class)) #the x axis will be the class of the car
pl + geom_bar(aes(fill=drv))
pl + geom_bar(aes(fill=drv),position = "dodge") 
pl + geom_bar(aes(fill=drv),position = "fill")#we make the color dependent of the drive type to make a stacked bar plot, cool!

#Dodge and fill are types of stacking of the bars
```

#**Boxplots**
TO graph the **quartiles as boxes**, on the x axis should be a categorical or discrete value (we need to use values)
- Bottom and top of the box are: 1st and 3rd quartiles  
- The band in the middle: 2nd quartile (median of the data)  
- The ends of the lines: 1.5x the interquartile range (upper and lower)

```{r}
df <-  mtcars
head(mtcars)

pl <- ggplot(df,aes(x=factor(cyl),y=mpg))

pl + geom_boxplot(aes(fill=factor(cyl))) + coord_flip()
```

#**2 Variables plotting**
```{r}
library(ggplot2)
library(ggplot2movies)

pl <- ggplot(movies,aes(x=year,y=rating))
pl2 <- pl + geom_bin2d(binwidth=c(3,1))
pl2 + scale_fill_gradient(low = "blue",high = "red")

#we can also use hexagons, if you wish
pl <- ggplot(movies,aes(x=year,y=rating))
pl2 <- pl + geom_hex()
pl2 + scale_fill_gradient(low = "blue",high = "red")

#Geom to density plot
pl <- ggplot(movies,aes(x=year,y=rating))
pl2 <- pl + geom_density2d_filled()
pl2

#can also be unfilled
pl <- ggplot(movies,aes(x=year,y=rating))
pl2 <- pl + geom_density2d()
pl2
```

#**Coordinates and faceting**
Coordinates help us to size correctly our plot
Faceting allows us to put more than one plot in the window
```{r}
library(ggplot2)

#Using coordinates
pl <- ggplot(mpg,aes(x=displ,y=hwy)) + geom_point()
pl + coord_cartesian(xlim = c(1,4),ylim = c(15,30))

print(pl)

#Using facets
pl <- ggplot(mpg,aes(x=displ,y=hwy)) + geom_point()
pl + facet_grid(. ~ cyl) #facet_grid(yaxis ~ xaxis)
pl + facet_grid(drv ~ .)
pl + facet_grid(drv ~ cyl)
```

#**Themes**
It gives beauty to our plots
```{r}
library(ggplot2)
library(ggthemes) #it has lots of options of themes

#theme_set(theme_minimal()) #the theme_set sets the theme for ALL the plots in the script otherwise do the below

pl <- ggplot(mtcars,aes(x=wt,y=mpg)) + geom_point() + theme_excel()

print(pl)

```

#**..ggplot2 Exercises..**
For the first few plots, use the mpg dataset

Histogram of hwy mpg values:
```{r}
df <- mpg
ggplot(df,aes(x=hwy)) + geom_histogram(fill = "red",alpha = 0.5, bins = 20) + theme_gray()
```
Barplot of car counts per manufacturer with color fill defined by cyl count
```{r}
df <- mpg
ggplot(df,aes(x=manufacturer)) + geom_bar(aes(fill=factor(cyl)))
```
Switch now to use the txhousing dataset that comes with ggplot2
```{r}
df <- txhousing
ggplot(df,aes(x=sales,y=volume)) + geom_point(aes(color=volume),alpha=0.5)
```

Add a smooth fit line to the scatterplot from above. Hint: You may need to look up geom_smooth()
```{r}
df <- txhousing
ggplot(df,aes(x=sales,y=volume)) + geom_point(aes(color=volume),alpha=0.5,size=5) + geom_smooth(color = "green")
```

