---
title: "Section 9 - R Data Frames"
author: "Victoria"
date: "3/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ***Data frames basics***
We can combine different data types in a dataframe, unlike matrices
***Data frames are similar to an Excel spreadsheet***
```{r}
state.x77 #Data frame with info of the states of the usa
state.area
USPersonalExpenditure #info of the US personal expenditure
women #Us women's height and weight
```
### To get the list of all the data frames (or datasets) available in R, you can do the following
```{r}
data()
```
### To show only some parts of the data frame
```{r}
head(state.x77) #it returns only the FIRST 6 rows of data
tail(state.x77) #it returns only the LAST 6 rows of data
str(state.x77) #it shows the attributes of the structure of the data frame
summary(state.x77) #it gives a summary of each of the categories of info in the data frame
```
## ***To create our own data frame***
```{r}
days <- c("Mon","Tue","Wed","Thur","Fri")
temp <- c(22.2,21,23,24.3,25)
rain <- c(T,T,F,F,T)

data.frame(days,temp,rain) #to create our data frame out of vectors!
df <- data.frame(days,temp,rain)
df

str(df)
summary(df)
```

# ***Data frames selection and indexing***
First we load the data frame
```{r}
days <- c("Mon","Tue","Wed","Thur","Fri")
temp <- c(22.2,21,23,24.3,25)
rain <- c(T,T,F,F,T)

data.frame(days,temp,rain) #to create our data frame out of vectors!
df <- data.frame(days,temp,rain)
df
```
Lets start indexing!
```{r}
df[1,] #all the columns of the 1st row
df[,1] #all the rows of the first column
df[,"days"]
df[,"rain"]
df[1:3,c("days","temp")] #rows 1 to 3 and only the columns "days" and "temp"
```
We can also use another quick notations to grab columns of a data frame as a **VECTOR**
```{r}
df$days #returns the column of days as a vector
df["days"] #returns the column of days as a data frame
```
Lets learn to use the **subset function**
```{r}
subset(df,subset = rain==TRUE) #it only returns the values in which rain==TRUE is true
subset(df,subset = temp>23)
```
We can also **order** a data frame
```{r}
sorted.temp <- order(df["temp"]) #it creates a vector of the ascending order of the temp column values
sorted.temp
df[sorted.temp,] #it returns the temp values in order from low to high

# Now in descending order
desc.temp <- order(-df["temp"])
desc.temp
df[desc.temp,]
```

# **Overview of Data Frame Operations**
## Creating Data Frames
```{r}
empty <- data.frame() #to create an empty data frame
c1 <- 1:10 #vector from 1 to 10
c1
c2 <- letters #vector of letters from a to z
c2

df <- data.frame(col.name.1 = c1, col.name.2 = c2[1:10]) #the col.name.n will be actually the name you want to give to your columns

```

##Working with **CSV files**
```{r}
#d2 <- read.csv("some_file.csv") #to read and assign the csv file to a variable
write.csv(df,file = "saved_df.csv") #to create a csv file out of a data frame we created here
df2 <- read.csv("saved_df.csv") #now, lets read it
df2
df
```

## Getting information about the data frame
```{r}
df
nrow(df) #it tells us how many rows we have
ncol(df) #it tells us how many columns we have (not including the index column)
colnames(df) #names of columns
rownames(df) #names of rows
summary(df)
str(df)
```

## Referencing cells
```{r}
df
df[[5,2]] # to reference the info in the row 5, column 2 in a data frame
df[[5,"col.name.2"]]
```
To **re-assign a value inside of a made data frame**
```{r}
df[[2,"col.name.1"]] <- 9999 #we changed the value in the referenced position
df
```
# Referencing rows
```{r}
df[1,]
```
# Referencing columns
```{r}
mtcars #an already made data frame
head(mtcars)

# To return a column as a VECTOR
mtcars$mpg 
mtcars[,"mpg"]
mtcars[,1]
mtcars[["mpg"]]

# To return a column as a DATA FRAME
mtcars["mpg"]
mtcars[1]

# To return multiple columns as a DATA FRAME
head(mtcars[c("mpg","cyl")])

```
## Adding rows
```{r}
df2 <- data.frame(col.name.1 = 2000,col.name.2 = "new") #we create a chunk of data frame that we want to add to the previous data frame df
df

dfnew <- rbind(df,df2) #we bind the new row
dfnew
```
## Adding columns
```{r}
df$newcol <- 2*df$col.name.1 # we add a new column with the value of 2 times the col.name.1 of df
df

df$newcol.copy <- df$newcol #we created another col
df
# we can also do this
df[,"newcol.copy2"] <- df$newcol
```

## Setting column names
```{r}
ncol(df)
colnames(df) <- c("col1","col2","col3","col4","col5")

# To name a specific column
colnames(df)[1] <- "new col name" #we just want to rename the 1st column
```

## Selecting multiple rows
Conditioning with $ symbol
```{r}
df[1:10,] #to select the rows 1 to 10
df[1:3,]
head(df,7) #to return the first 7 rows
df[-2,] #to select EVERYTHING BUT THE 2ND ROW

#Filtering the rows that we want to get
head(mtcars)
mtcars[mtcars$mpg > 20,] #to return only therows in which mpg value is greather than 20
mtcars[(mtcars$mpg>20) & (mtcars$cyl == 6),c("mpg","cyl","hp")] #to get only the selected filtered rows and columns
```
Conditioning with **Subset function**
```{r}
subset(mtcars, mpg>20 & cyl == 6)
```

## Selecting multiple columns
```{r}
head(mtcars)
mtcars[,c(1,2,3)]
mtcars[,c("mpg","cyl")]
```

## Dealing with missing data
Missing data = NA :(
```{r}
is.na(mtcars) #it returns FALSE if nothing is missing
any(is.na(mtcars)) #it checks QUICKLY if there's any TRUE in the NA check of all the data frame
any(is.na(mtcars$mpg))
```
TO replace missing data
```{r}
df[is.na(df)] <- 0
mtcars$mpg[is.na(mtcars$mpg)] <- mean(mtcars$mpg) #only replaces the missing values of a column
```

#**...Data frames exercises...**
Ex 1: Recreate the following dataframe by creating vectors and using the data.frame function:

```{r}
name <- c("Sam","Frank","Amy")
age <- c(22,25,26)
weight <- c(150,165,120)
sex <- c("M","M","F")

adults.info <- data.frame(name,age,weight,sex)
adults.info
```
Ex 2: Check if mtcars is a dataframe using is.data.frame()
```{r}
is.data.frame(mtcars)
```
Ex 3: Use as.data.frame() to convert a matrix into a dataframe:
```{r}
mat <- matrix(1:25, nrow = 5)
mat.df <- as.data.frame(mat)
```
Ex 4: Set the built-in data frame mtcars as a variable df. We'll use this df variable for the rest of the exercises.
```{r}
df <- mtcars
```
Ex 5: Display the first 6 rows of df
```{r}
head(df)
```
Ex 6: What is the average mpg value for all the cars?
```{r}
mean(mtcars$mpg)
```
Ex 7: Select the rows where all cars have 6 cylinders (cyl column)
```{r}
# 2 ways of doing it
subset(mtcars, mtcars$cyl == 6)
mtcars[mtcars$cyl == 6,]
```
Ex 8: Select the columns am,gear, and carb.
```{r}
mtcars[,c("am","gear","carb")]
```
Ex 9: Create a new column called performance, which is calculated by hp/wt.

```{r}
mtcars$performance <- mtcars$hp / mtcars$wt
mtcars
head(mtcars)
```
Ex 10: Your performance column will have several decimal place precision. Figure out how to use round() (check help(round)) to reduce this accuracy to only 2 decimal places.
```{r}
mtcars$performance <- round(mtcars$hp / mtcars$wt, digits = 2)
mtcars
head(mtcars)
```
Ex 10: What is the average mpg for cars that have more than 100 hp AND a wt value of more than 2.5.
```{r}
mean(mtcars[mtcars$hp > 100 & mtcars$wt > 2.5,"mpg"])
```
Ex 11: What is the mpg of the Hornet Sportabout?
```{r}
mtcars["Hornet Sportabout","mpg"]
```

