---
title: 'Section 14: Data Manipulation with R'
author: "Victoria"
date: "7/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**Guide to using Dplyr**
```{r}
 #dataset we're gonna play with
library(dplyr)
library(nycflights13)
```
Lets start playing with the data
```{r}
head(flights)
summary(flights)
```
##How to use dplyr to easily manipulate our data
FUnctions inside of dplyr to use
###Filter():to select a set of rows in a df
```{r}
head(filter(flights,month==11,day==3,carrier=="AA")) #with "filter" function

```
###Slice(): to select rows by position
```{r}
slice(flights,1:10) #we select the row 1 to 10
```
###arrange(): re-orders the rows of a df in ascending order
```{r}
head(arrange(flights,year,month,day,arr_time)) #it orders the rows in the written columns in ascending order
head(arrange(flights,year,month,day,desc(arr_time)))
```
###select(): allows to only show specific columns in df
```{r}
head(select(flights,carrier))
head(select(flights,carrier,dep_time))
```
##rename(): it lets us rename the columns of a df
```{r}
head(rename(flights,airline_carrier = carrier)) #new name = old name
```
##distinct(): allows us to select the unique values in a column or table
```{r}
distinct(select(flights,carrier)) #it only shows the different unique values (doesn't show the repetition)
```
##mutate(): to create new columns
```{r}
head(mutate(flights,new_col = arr_delay - dep_delay))
```
##transmute(): it creates a new column and ONLY returns this new column
```{r}
head(transmute(flights,new_col = arr_delay - dep_delay))
```
##summarize(): allows us to collapse dataframes in single rows
```{r}
summarise(flights,avg_air_time = mean(air_time,na.rm = TRUE)) #it returns one result as a row
summarise(flights,total_air_time = sum(air_time,na.rm = TRUE))
```

##sample()
```{r}
sample_n(flights,10) #returns 10 random rows of the df
sample_frac(flights,0.1) #returns a fraction of the rows (in this case it returns the 10% of the rows)
```

# **Pipe Operator %>%**
It allows us to make sequential operations like: data %>% op1 %>% op2...
```{r}
library(dplyr)

df <- mtcars

#Nesting
result <- arrange(sample_n(filter(df,mpg>20),size=5),desc(mpg))
print(result)

#the same but using Pipe operator
result.pipe <- df %>% filter(mpg>20) %>% sample_n(size=5) %>% arrange(desc(mpg))

print(result.pipe)
```

#**..Dplyr exercises..**
```{r}
library(dplyr)
head(mtcars)
```
Return rows of cars that have an mpg value greater than 20 and 6 cylinders.
```{r}
res <- filter(mtcars,mpg>20,cyl==6)
head(res)
```

Reorder the Data Frame by cyl first, then by descending wt.
```{r}
res <- arrange(mtcars,cyl,desc(wt))
head(res)
```

Select the columns mpg and hp
```{r}
res <- select(mtcars,mpg,hp)
head(res)
```

Select the distinct values of the gear column.
```{r}
distinct(select(mtcars,gear))
```

Create a new column called "Performance" which is calculated by hp divided by wt.
```{r}
res <- mutate(mtcars, Performance = hp/wt)
head(res)
```

Find the mean mpg value using dplyr.
```{r}
summarize(mtcars,avg_mpg = mean(mpg,na.rm = TRUE))
```

Use pipe operators to get the mean hp value for cars with 6 cylinders.
```{r}
res <- mtcars %>% filter(cyl==6) %>% summarise(std_hp = mean(hp,na.rm = TRUE))
print(res)
```

#**Tidyr**
It's a function that helps us to clean data, this means, that each column is a property or variable, and each row is a vehicle with values
##Data table = Data frame, but the data table has some extra features, and it's faster to process than a df
```{r}
library(tidyr)
library(data.table)
```
##gather(): it gathers different columns into one single column
```{r}
#First, we create fake data to process it
comp <- c(1,1,1,2,2,2,3,3,3)
yr <- c(1998,1999,2000,1998,1999,2000,1998,1999,2000)
q1 <- runif(9, min=0, max=100) # random 9 numbers from 0 to 100
q2 <- runif(9, min=0, max=100)
q3 <- runif(9, min=0, max=100)
q4 <- runif(9, min=0, max=100)

df <- data.frame(comp=comp,year=yr,Qtr1 = q1,Qtr2 = q2,Qtr3 = q3,Qtr4 = q4)

gather(df,Quarter,Revenue,Qtr1:Qtr4) #we gather all the information of the Qtr1,2,3... into 2 columns: key = Quarter and value = Revenue

df %>% gather(Quarter,Revenue,Qtr1:Qtr4) #it does the same
```
## spread(): it's a opposite of gather
```{r}
stocks <- data.frame(
  time = as.Date('2009-01-01') + 0:9,
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4)
)
stocks
#lets use GATHER()
stocks.gathered <- stocks %>% gather(stock,price,X,Y,Z) #key = stock and value = price (X,Y,Z)
head(stocks.gathered)

#lets use SPREAD()
stocks.gathered %>% spread(stock,price) #we spread putting (key,values)
stocks.gathered %>% spread(time,price) #we can also do this, but it is not really tidy

```
## separate(): it transforms a single character column into multiple columns
```{r}
df <- data.frame(new.col=c(NA,"a.x","b.y","c.z"))
df

separate(df,new.col,c("ABC","XYZ")) #it separates non alphanumerical elements to create columns (like - , . / etc...)
```
## unite(): the opposite of separate
```{r}
df <- data.frame(new.col=c(NA,"a.x","b.y","c.z"))
df.sep <- separate(df,new.col,c("ABC","XYZ")) 

unite(df.sep, new.joined.col, ABC, XYZ, sep = "---") #it unites the columns ABC and XYZ as a new column new.joined.col
```

